---
title: "Basic Tutorial"
author: "Dana Gibbon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Quick start

* Install `PressPurtCoreAlg`

```{r install, eval=FALSE}
# Specify branch and build vignettes
devtools::install_github("dkoslicki/PressPurtCoreAlg", 
                         ref = "gibbond", build_vignettes = TRUE)
```

* Load the library
* Check available python versions
* Set python with previously made conda environment
* install python dependencies
* Run PreprocessMatrix
* Run ComputeEntryWisePerturbationExpectation
* Plot
    * specific figures
    * all figures

```{r quick_start_setup}
library(PressPurtCoreAlg)
find_python()
set_python("r-reticulate", verbose = TRUE)
```

```{r quick_start_install, eval=FALSE}
py_depend("r-reticulate")
```

```{r quick_start_run}
# Path to your matrix
infile <- system.file("extdata", "Modules", "IGP.csv", package = "PressPurtCoreAlg")
MultiPert <- ComputeMultiEntryPerturbationExpectation(input_file = infile)
PreProsMatrix <- PreprocessMatrix(input_file = infile, 
                                  output_folder = NULL, 
                                  max_bound = 10, 
                                  threads = 2)
Entrywise <- ComputeEntryWisePerturbationExpectation(PreProsMatrix = PreProsMatrix, 
                                        distribution_type="truncnorm", 
                                        input_a=0, input_b=-2, threads=1)
list_of_numswitch_to_plot <- list(c(1, 1), c(1, 2))
GenerateEntryWiseFigures(EntryWise=Entrywise, 
                         all_numswitch_plots = FALSE, 
                         list_of_numswitch_to_plot=list_of_numswitch_to_plot)
```

```{r all_plots, fig.width=9, fig.height=8}
GenerateEntryWiseFigures(EntryWise=Entrywise, 
                         all_numswitch_plots = TRUE)
```

# Load Python

This package uses python through the R package `reticulate`. You'll need to install the needed python libraries into a conda env or a virtualenv. The "dependencies_tutorial" walks you through how to do this. Once this is done, you'll need to load your conda env or virtualenv.

Note: If you want to change conda envs, you'll need to restart R because of how `reticulate` works.

```{r load_py, eval=FALSE}
# set the conda env
set_python("r-reticulate", verbose = TRUE)
```

# Example: IGP.csv
When running this workflow, one may keep everything in R memory or save the output and intermediate files to a specified folder. However, many of these output files are python objects. Let's start by looking at the R-based version:

## R-object based

### Compute Multi Entry Perturbation Expectation
This function takes a jacobian matrix and computes the multi-entry perturbation expectation.

In this example we take the `IGP.csv` matrix and use the default `num_iterates` and `interval_length` values.

```{r multi}
infile <- system.file("extdata", "Modules", "IGP.csv", package = "PressPurtCoreAlg")
MultiPert <- ComputeMultiEntryPerturbationExpectation(input_file = infile)
MultiPert
```

### PreprocessMatrix

This function pre-processes a matrix by figuring out what the intervals of asymptotic stability are, as well as finding which perturbation values lead to a sign switch.

In this example we use a `max_bound` of 10, don't perturb zero entries and use 2 threads.

```{r PrePros}
PreProsMatrix <- PreprocessMatrix(input_file = infile,
                                  max_bound = 10,
                                  zero_perturb = FALSE,
                                  threads = 2)
```

#### Output

The output is a list of 9 objects which includes:
`r names(PreProsMatrix)`

##### Matrix

This includes the original matrix and information about that matrix:
```{r matrix}
# Original Matrix
PreProsMatrix$original_matrix
# Matrix size
PreProsMatrix$matrix_size
# row names
PreProsMatrix$row_names
# column names
PreProsMatrix$column_names
# number of entries that aren't zero
PreProsMatrix$non_zero
```

##### Asymptotic of Stability intervals

The asymptotic of stability intervals are divided up into two matrices: start and stop.

```{r AS}
# Start
PreProsMatrix$asymptotic_stability_start
# End
PreProsMatrix$asymptotic_stability_end
# To access entry 1,1:
PreProsMatrix$asymptotic_stability_start[1,1]
PreProsMatrix$asymptotic_stability_end[1,1]
```

##### Num Switch

The `num_switch_funcs` object is a list of each non-zero entry in the matrix and the corresponding perturbation values that leads to a sign switch.

For example, entry 1,1 has 1 sign switch and 2,2 has 2.

```{r num_switch}
# get the length of the list, see it's 
# the same as the non-zero entries
length(PreProsMatrix$num_switch_funcs_r)
# see the names of the entries
names(PreProsMatrix$num_switch_funcs_r)
# num switch function for 1,1
PreProsMatrix$num_switch_funcs_r$`(1, 1)`
# num switch function for 2,2
PreProsMatrix$num_switch_funcs_r$`(2, 2)`
```

You can ignore the `num_switch_functions_py` object as it is a python dictionary that is passed to the next step.

### Compute Entry Wise Perturbation Expectation

This function computes the expected number of sign switches from perturbing each entry individually.

In this example, the `PreProsMatrix` object is passed, "truncnorm" distribution is used, `input_a` is 0 and `input_b` is -2.

```{r entrywise}
Entrywise <- ComputeEntryWisePerturbationExpectation(PreProsMatrix = PreProsMatrix, 
                                        distribution_type="truncnorm", 
                                        input_a=0, input_b=-2, threads=1)
```

#### Output

The output includes the same object that were found in `PreProsMatrix` with 3 additions: 
`r names(Entrywise)`

##### Expected num switch

This is a data frame of expected sign switches.

```{r exp_NS}
Entrywise$expected_num_switch
```

##### Distributions Object

This object gets the probability density distribution into a plot ready format for each non-zero interval in the matrix. By default there are 250 points in the distribution.

```{r dists_ob}
# distributions object of each non-zero entry in the matrix
names(Entrywise$distributions_object)
# position in the matrix
Entrywise$distributions_object$`(1, 1)`[1]
# Interval of asymptotic of stability
Entrywise$distributions_object$`(1, 1)`[2]
# ready to plot probability density distribution dataframe
head(Entrywise$distributions_object$`(1, 1)`[[3]], n = 10)
```

##### Num Switch Object

The `ns_object_plot` is the Num Switch functions that have been converted into step functions and are ready to plot!

```{r NS_plot}
# Step function for each non-zero entry in the matrix
names(Entrywise$ns_object_plot)
# entry 1,1
Entrywise$ns_object_plot$`(1, 1)`
# entry 4,2
Entrywise$ns_object_plot$`(4, 2)`
```

The y-0 position has been added to the step function:

```{r compare_NS}
# Before
Entrywise$num_switch_funcs_r$`(1, 1)`
# After - step function
Entrywise$ns_object_plot$`(1, 1)`
```

### Generate Entry Wise Figures

The `GenerateEntryWiseFigures` function plots the Num Switch step function on the Y-axis in black and the probability density on the Y-axis in grey. 

One may plot individual entries or a list of entries with the `list_of_numswitch_to_plot` option:

```{r plot_entry}
list_of_numswitch_to_plot <- list(c(1, 1), c(1, 2))
GenerateEntryWiseFigures(EntryWise=Entrywise, 
                         all_numswitch_plots = FALSE, 
                         list_of_numswitch_to_plot=list_of_numswitch_to_plot)

```

Or plot all figures with `all_numswitch_plots`. 

For large matrices it is not recommended that you plot them all in one plot.

```{r plot_all, fig.width=9, fig.height=8}
GenerateEntryWiseFigures(EntryWise=Entrywise, 
                         all_numswitch_plots = TRUE)
```

## Save to disk

### Basic pipeline

One may instead save objects to disk if one sets an `output_folder`. The following files will be created in the specified folder:

* asymptotic_stability.npy
* distributions.pkl
* num_non_zero.npy
* row_names.txt
* column_names.txt
* expected_num_switch.csv  
* num_switch_funcs.pkl
* size.npy

If you are going to be doing multiple runs in the same folder it is recommended that you specify a file prefix - this will add <prefix>_ to the beginning of every file.

```{r save_disk, eval=FALSE}
# Not run in vignette creation
PreProsMatrix <- PreprocessMatrix(input_file = infile, 
                                  output_folder = "test_r/test4", 
                                  prefix = "IGP", max_bound = 10, threads = 2)
Entrywise <- ComputeEntryWisePerturbationExpectation(input_folder = "test_r/test4",
                                                     prefix = "IGP",
                                                     distribution_type="truncnorm", 
                                                     input_a=0, input_b=-2, threads=1)

list_of_numswitch_to_plot <- list(c(1, 1), c(1, 2))
GenerateEntryWiseFigures(input_folder = "test_r/test4",
                         prefix = "IGP",
                         all_numswitch_plots = FALSE, 
                         list_of_numswitch_to_plot=list_of_numswitch_to_plot)

```


Many of them are numpy or pickle python objects. But the `GenerateEntryWiseFigures` can handle them!

### Process Data

One may also turn the data saved to a folder into an R friendly format with `process_data`. This function will put the data in the same format as the output as `ComputeEntryWisePerturbationExpectation`.

```{r proc_disk, eval=FALSE}
Entrywise_from_disk <- process_data(matrix = infile, 
                                    type = "csv", 
                                    folder = "test_r/test4", 
                                    prefix = "IGP")
```

