---
title: "Set up"
author: "Dana Gibbon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Installing Dependencies

## R

PressPurtCoreAlg depends on R >= 3.1.0. To check your R version run `version` from the R console.

This package also depends on the following R libraries:

* `data.table` 
* `ggplot2`
* `gridExtra`
* `reticulate >= 1.11`

If not already installed, these packages will be installed upon the installation of `PressPurtCoreAlg`.

## Python

This package is dependent on python and uses the R package `reticulate` under the hood to use python. Thus python and the dependant packages must be installed. 

There are two suggested approaches: 

* Conda (Anaconda/miniconda) which supports Linux/Mac/Windows
* Virtual environments which supports Linux/Mac OS

It is also possible to set a specific version of python with `reticulate::use_python("/path/to/python")` before you run any of the `PressPurtCoreAlg` commands but it is not recommended. You'll also have to ensure that the dependent python packages are installed. For more information on using `reticulate` directly: [see the reticulate docs](https://rstudio.github.io/reticulate/articles/python_packages.html).

### Conda

#### Install Anaconda or Miniconda 

Conda is a package and environment manager that is open source. The main difference between Anaconda and Miniconda is that Anaconda comes with a bundle of pre-installed packages so it takes longer to download and install. Both Miniconda and Anaconda come with python but you can specify a specific version of python later as well.

This document will show you how to install Miniconda via the command-line. For more information on installation and how to install it graphically (no command-line) [see the conda docs](https://docs.conda.io/projects/continuumio-conda/en/latest/user-guide/install/index.html)

##### Linux

1. [Download the Miniconda installer for Linux](https://docs.conda.io/en/latest/miniconda.html)
2. In a terminal window run: `bash Miniconda3-latest-Linux-x86_64.sh`
3. Follow the prompts - accepting defaults should be fine.
4. Close your Terminal window and re-open a new one.
5. Test your installation by running: `conda list`
    * This should result in a list of installed packages.

##### MacOS

1. [Download the Miniconda installer for macOS](https://docs.conda.io/en/latest/miniconda.html)
2. In a terminal window run: `bash Miniconda3-latest-MacOSX-x86_64.sh`
3. Follow the prompts - accepting defaults should be fine.
4. Close your Terminal window and re-open a new one.
5. Test your installation by running: `conda list`
    * This should result in a list of installed packages.

##### Windows

1. [Download the Miniconda installer for Windows](https://docs.conda.io/en/latest/miniconda.html)
2. Run the `.exe` file by double clicking.
3. Follow the prompts - accepting defaults should be fine.
5. Test your installation by opeing the Anacond Prompt from the Start menu and running running: `conda list`
    * This should result in a list of installed packages.

#### Installing python dependencies

##### From R

First you'll need to load `PressPurtCoreAlg`, check which versions of python are found and if any Conda envrionments exists.

```{r}
# load the library
library(PressPurtCoreAlg)
# Find the versions of python and conda environments available on your machine:
find_python()

```

This shows you your default Python path, other available python paths, and a list of your conda environments.

The `set_python` function lets one specify a conda envrionemnt to use. If you want to make a new conda environment, you'll need to specify which python version to use in addition to the conda envirnment name.

You may have python installed under `/usr/bin/python3`, Anaconda2 and/or Anaconda3 installed so it's important to specify which python to use. In this case I'll use the anaconda3 python.

```{r, eval=FALSE}
set_python(condaenv="test-r", version="/home/gibbond/anaconda3/bin/python",verbose = TRUE)
```

Next you'll need to install the python dependencies in your conda environment. This will install the following into your `test-r` conda env: 

* `numpy`
* `scipy`
* `matplotlib`
* `sympy`
* `pathos`
* `pandas`

```{r, eval=FALSE}
py_depend("test-r")
```

If the dependencies don't install correctly see below for instructions on how to install via the command-line.

If you want to use the `test-r` environment in a new R session you can access it via:

```{r, eval=FALSE}
set_python(condaenv="test-r")
```

##### Command-line

You can also install your Python dependencies via the command line with conda or with the Anaconda Prompt (Windows).

```{bash, eval=FALSE}
# Check that conda is running
conda --version
# List conda envs
conda info --envs
# Create conda environment "test-r"
conda create --name test-r
# activate test-r
source activate test-r
# check installed packages
conda list
# install python dependencies
conda install matplotlib numpy pandas scipy sympy
pip install pathos
# Check to make sure they were installed
conda list
```

Now that you have created the conda environments and installed dependencies, you can use the new env in R! Run:

```{r, eval=FALSE}
# load the library
library(PressPurtCoreAlg)
# You should see "test-r" listed as a conda env
find_python()
# Set your conda env
set_python(condaenv="test-r")
```

### Virtualenv


